---
title: "Data wrangling with sparklyr"
output: pdf_document
---

**Week 3 tutorial task**

Welcome to this tutorial paired-programming practice document! 

In pairs, work through this document. A sample completed document is available on the tutorial github repository as well.


# Task 1 

Install the following packages if you do not already have them already, and load them with the `library()` function: 

* sparklyr 
* dplyr 
* ggplot2
* knitr 

```{r warning=FALSE, message=FALSE}
library(sparklyr)
library(dplyr)
library(ggplot2)
library(knitr)
```


# Task 2

Create a spark connection that treats your local machine as a cluster. Load up the msleep dataset from the ggplot2 package, and put it into a Spark dataframe.

```{r}
sc <- spark_connect(master = "local")
msleep_spark <- copy_to(sc, msleep, overwrite = TRUE)
```


# Task 3

Which dplyr functions can be used to have a first look at the data? Run them.

```{r}
# dplyr functions
msleep_spark %>%
  head()

msleep_spark %>%
  glimpse()

msleep_spark %>%
  str()

msleep_spark %>%
  sdf_describe()

msleep_spark %>%
  colnames()

msleep_spark %>%
  head() %>%
  kable()
```


# Task 4

Count the occurrence of all the different values for the order variable that are
present in the dataset. Use kable to display any tables.

```{r}
msleep_spark %>%
  count(order) %>%
  kable()

```

# Task 5

Select the columns name, order and brainwt. Filter to only those rows where
the order variable is Primates. Finally, arrange the result by brainwt. 
Try to use the pipe operator %>% to cut down on how many lines of code you write.

```{r}
msleep_spark %>%
  select(name, order, brainwt) %>%
  filter(order=="Primates") %>%
  arrange(desc(brainwt)) %>%
  kable()
```

# Task 6

In this task, we will create a table of mean brain weight to body weight ratio,
grouped by the variable order.
Create a new column called brain_body_wt_ratio that is equal to brainwt/bodywt.
Group by the variable order, and then calculate the mean for each group.

```{r}
msleep_spark %>%
  mutate(braint_body_wt_ratio = brainwt / bodywt) %>%
  group_by(order) %>%
  summarise(mean_order = mean(braint_body_wt_ratio, na.rm = TRUE)) %>%
  arrange(desc(mean_order))
```

# Task 7

The total number of missing values in each column can be viewed using the following code

```{r warning = FALSE}
msleep_spark %>%
  summarise_all(~sum(as.integer(is.na(.)))) %>%
  kable()
```

Impute missing values for brainwt. Do this by creating a new column called
brainwt_imputed, where NA values are replaced with the mean value for brainwt.
Verify the result by displaying the head of a table with the columns name, 
brainwt and brainwt_imputed. You may find the ifelse or case_when functions
useful (Google them).

```{r}
msleep_spark %>%
  mutate(brainwt_imputed = ifelse(is.na(brainwt),
                                  mean(brainwt, na.rm = TRUE),
                                  brainwt)) %>%
  select(name, brainwt, brainwt_imputed) %>%
  head()
```


# Task 8

Use collect() and ggplot to make a horizontal bar chart with order on the vertical axis and average sleep_total on the horizontal axis.
You may find geom_col from ggplot useful.

```{r}
sleep_order <- msleep_spark %>% 
   group_by(order) %>% 
   summarise(mean = mean(sleep_total, na.rm = TRUE)) %>% 
   collect()

sleep_order %>% 
   ggplot(aes(x=order, y=mean)) +
   geom_col(fill="Steelblue") +
   xlab("Order")+
   ylab("Average Sleep Total") +
   labs(title="The Average Total Sleep for each order of Spieces") +
   coord_flip()
```
