---
title: "Introduction to sparklyr"
output: pdf_document
---

We will largely follow chapters **2** and **3** of **Mastering Spark with R**, [https://therinspark.com](https://therinspark.com/index.html).

First install the following packages if you do not already have them already, and load them with the `library()` function: 

```{r warning=FALSE, message=FALSE}
library(sparklyr)
library(dplyr)
library(ggplot2)
library(knitr)
```


## Preliminiaries

If you are working on EIDF, first make sure that the default working directory in RStudio is your folder. In RStudio, select Tools -> Global Options. Change the default working directory to be /work/eidf071/eidf071/<your username>.

To confirm the change has taken effect, close and then reopen RStudio, and type getpw() into the console. It should show your working directory correctly as above.

## Connecting

```{r}
sc = spark_connect(master = 'local')
```


# Task 3

Which dplyr functions can be used to have a first look at the data? Run them.

```{r}
#install.packages("nycflights13", "Lahman")
library(dplyr)
iris_tbl <- copy_to(sc, iris, OVERWRITE = TRUE)
flights_tbl <- copy_to(sc, nycflights13::flights, "flights", OVERWRITE = TRUE)
batting_tbl <- copy_to(sc, Lahman::Batting, "batting", OVERWRITE = TRUE)
src_tbls(sc)

flights_tbl %>% filter(dep_delay ==2)

delay <- flights_tbl %>% 
  group_by(tailnum) %>%
  summarise(count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE)) %>%
  filter(count > 20, dist < 2000, !is.na(delay)) %>%
  collect()

library(ggplot2)
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area(max_size = 2)
```


# Task 4

Count the occurrence of all the different values for the order variable that are
present in the dataset. Use kable to display any tables.

```{r}
# your code here
```

# Task 5

Select the columns name, order and brainwt. Filter to only those rows where
the order variable is Primates. Finally, arrange the result by brainwt. 
Try to use the pipe operator %>% to cut down on how many lines of code you write.

```{r}
# your code here
```


# Task 6

In this task, we will create a table of mean brain weight to body weight ratio,
grouped by the variable order.
Create a new column called brain_body_wt_ratio that is equal to brainwt/bodywt.
Group by the variable order, and then calculate the mean for each group.

```{r}
# your code here
```


# Task 7

The total number of missing values in each column can be viewed using the following code

```{r warning = FALSE}
sleep %>%
  summarise_all(~sum(as.integer(is.na(.)))) %>%
  kable()
```

Impute missing values for brainwt. Do this by creating a new column called
brainwt_imputed, where NA values are replaced with the mean value for brainwt.
Verify the result by displaying the head of a table with the columns name, 
brainwt and brainwt_imputed. You may find the ifelse or case_when functions
useful (Google them).

```{r}
# your code here
```


# Task 8

Use collect() and ggplot to make a horizontal bar chart with order on the vertical axis and average sleep_total on the horizontal axis.
You may find geom_col from ggplot useful.

```{r}
# your code here
```

