---
title: "Statistical modelling with sparklyr"
output: pdf_document
---

**Week 4 tutorial task**

In this exercise, we will be exploring some statistical models with sparklyr.

Load up the required packages.

```{r warning=FALSE, message=FALSE}
library(sparklyr)
library(dplyr)
library(ggplot2)
library(knitr)
library(broom)
```

Create a spark connection that treats your local machine as a cluster. Load up the built-in R dataset mtcars into a Spark dataframe.

```{r warning = FALSE}
sc <- spark_connect(master = "local")
cars <- copy_to(sc, mtcars, overwrite = TRUE)
```

# Task 1

One-hot encode the cyl variable. You may find the ifelse function
useful (Google it).

```{r}
cars_cyl <- cars %>% 
  mutate(cyl_4 = ifelse(cyl == 4, 1, 0),
         cyl_6 = ifelse(cyl == 6, 1, 0),
         cyl_8 = ifelse(cyl == 8, 1, 0))

# Show the result
cars_cyl %>%
  select(cyl, cyl_4, cyl_6, cyl_8) %>%
  head() %>%
  collect()
```


# Task 2

Create an OLS model that predicts the variable mpg using hp, wt and number
of cylinders as predictor variables. **Remember that categorical variables used as 
predictors must be one-hot encoded! Also, don't fall into the dummy variable trap!** 
Apply the tidy function from the broom package to the fitted model and display 
the results in a table. What do the entries in the table mean? What is their interpretation?

```{r}
# Fit the OLS model witho fit_intercept to prevent the dummy variable trap
ols_model <- cars_cyl %>%
  ml_linear_regression(mpg ~ hp + wt + cyl_4 + cyl_6 + cyl_8,
                       fit_intercept = FALSE)

# Apply the tidy function to the fitted model
model_summary <- tidy(ols_model)

# Display the results in a table
model_summary %>%
  collect() %>%
  kable()
```

# Task 3

The variable mpg **must** be positive. The exponential function has the property 
that it is always positive for any real value of its argument. We can take advantage
of this to create an OLS model where the dependent variable can never be negative,
regardless of the value of the coefficients or independent variables. This is 
done by replacing mpg with its logarithm in the model.

Create a new column in the Spark dataframe that is the logarithm of mpg. Then
fit a new OLS model with log(mpg) as the dependent variable. Display the
results in a table.

```{r}
cars_cyl <- cars_cyl %>%
  mutate(ln_mpg = log(mpg))

ols_model <- cars_cyl %>%
  ml_linear_regression(ln_mpg ~ hp + wt + cyl_4 + cyl_6 + cyl_8,
                       fit_intercept = FALSE)

#ols_model <- tidy(ols_model)

#ols_model %>%
#  collect() %>%
#  kable()
```


# Task 4

Recall that the upper/lower 95% confidence interval for fitted coefficients 
in a regression are given by adding/subtracting 1.96 standard deviations.

In the results table, create two new columns for the upper and lower limits
of the 95% confidence interval. Then create a column
that has the estimated coefficient, followed by a bracket with the 95% confidence interval, 
with all numbers rounded to 2 decimal places. You may find the paste and round functions
useful (Google them). Display the results in a table.

```{r}
ols_model_ci <- tidy(ols_model) %>%
  mutate(upper_ci = estimate + 1.96 * std.error,
         lower_ci = estimate + 1.96 * std.error) %>%
  mutate(ci = paste0(
    round(estimate, 2),
    " [", round(lower_ci, 2),
    ", ", round(upper_ci, 2), "]"
  )
  )

ols_model_ci %>%
  collect() %>%
  kable()
```


# Task 5

Evaluate the OLS model with ln(mpg) as the dependent variable using the ml_evaluate function. The main metrics that are
of interest are R^2, mean absolute error, mean squared error, and root mean 
squared error. Print these metrics out.

```{r}
ols_model_metrics <- ml_evaluate(ols_model, cars_cyl)
ols_model_metrics$r2
ols_model_metrics$mean_absolute_error
ols_model_metrics$mean_squared_error
ols_model_metrics$root_mean_squared_error
```


# Task 6

Calculate the mean squared error for your model predictions "by hand". Remember to use dplyr functions only!

```{r}
predictions <- ml_predict(ols_model, cars_cyl)
predictions

values <- mutate(predictions,
  squared_error = (ln_mpg - prediction)^2
) %>%
  summarise(mse = mean(squared_error, na.rm = TRUE)) %>%
  pull()

identical(values, ols_model_metrics$mean_squared_error)
```

